## ```searchAQ```

```searchAQ``` is an R package that includes utility to read in SEARCH air quality network data
in its original format and combine it into consistent lists and data frames. Other included 
functions perform detrending and calculate photochemical state from observations.

## Getting started
Download all SEARCH data from Dropbox, using one of these two links:
1) https://www.dropbox.com/sh/o9hxoa4wlo97zpe/AACbm6LetQowrpUgX4vUxnoDa?dl=0
2) https://www.dropbox.com/sh/1tsdqdmiwixbgfd/AACxwbKilNVbC52H1VH4geJma?dl=0

Maintain the original structure, with directories named for each year.

### Install ```searchAQ```
In R or RStudio, install the ```searchAQ``` R package from github.  

#### Install and load ```devtools```
```
install.packages( c('devtools'))
library(devtools)
```

#### Install and load ```searchAQ```
```{r}
devtools::install_github("lhenneman/searchAQ")
library(searchAQ)
```

## Reading the data
searchAQ relies on the original file formats for SEARCH data, which change 
in different years (2008, for example, saw a big change in file format).
For Jefferson Street, for example:
```{r}
JSTdata <- read_SEARCH('JST',loc.search)
```
Where ```loc.serach``` is the local path of your SEARCH data.
```JSTdata``` will have the structure of a list containing hourly
gas, meteorology, and ion measurements, and daily PM2.5:

```{r}
summary(JSTdata)
#     Length Class      Mode
# gas 9      data.frame list
# met 8      data.frame list
# pm  2      data.frame list
# ion 4      data.frame list
summary(JSTdata$gas[, 1:4])
#      date                           o3                co              so2         
# Min.   :1998-07-31 20:00:00   Min.   : -3.487   Min.   :-829.0   Min.   : -0.766  
# 1st Qu.:2002-09-08 01:45:00   1st Qu.:  8.153   1st Qu.: 192.8   1st Qu.:  0.490  
# Median :2006-10-16 07:30:00   Median : 21.627   Median : 254.2   Median :  1.381  
# Mean   :2006-10-16 07:30:00   Mean   : 25.019   Mean   : 377.5   Mean   :  3.533  
# 3rd Qu.:2010-11-23 12:15:00   3rd Qu.: 36.582   3rd Qu.: 380.5   3rd Qu.:  3.710  
# Max.   :2014-12-31 18:00:00   Max.   :167.186   Max.   :7687.5   Max.   :129.454  
#                               NA's   :5848      NA's   :8651     NA's   :8922     

```

## Creating daily metrics





